\artus -- Framework for Event-based Data Analysis in High Energy Physics

### 2015-10-14 Core/doc/latex/paper.tex
Paper: First easy to fix issues we discussed

\begin{abstract}
We present \artus, an event-based data-processing framework for high energy physics experiments.
\mbox{\artus} is designed for large-scale data analysis in a collaborative environment.
The architecture design choices take into account typical challenges and are based on experiences with similar applications.
The structure of the framework and its advantages are described.
An example use case and performance measurements are presented.
The framework is well-tested and successfully used by several analysis groups.\todo[inline]{Mention them, maybe with publications? -> mentioned at the beginning of Htt chapter. Do we have publications?}
\end{abstract}

@@

\section{Common Workflow of High Energy Physics Analyses}
\label{section_artus_analysis_workflow}
% This section should contain a short description of a HEP/CMS-typical workflow.
% Especially common problems and challenges should be highlighted.
% From these problems, the design requirements for Artus are derived.
The workflow of the experiments collecting data at the Large Hadron Collider (LHC) at CERN is taken as an example to illustrate the organization and needs of a modern High Energy Physics (HEP) analysis.
Still, equal or very similar requirements are also present for HEP experiments not located at the LHC.

Between the data taking at the detector or the generation of simulated events with the subsequent detector simulation and the showcasing of the physical results, the following analysis steps have shown to be the best compromise in terms of turnaround time and efficiency as explained in Figure \ref{figure_analysis_steps}.

The reconstruction of the raw data coming from the experiments is performed on large computing centers organized into the WLCG\footnote{\url{http://wlcg-public.web.cern.ch}}, the Worldwide LHC Computing Grid.
The resulting datasets contain event information and high-level physics objects and are of $\mathcal O(10-100)$~TB in size.
\todo[inline]{move file size numbers to Figure \ref{figure_analysis_steps}?}

@@

Source code and configuration files have become the most precise form of documentation.
Therefore, readability also for people who have not written the code is crucial.
Also, if possible, existing and well-established libraries of the HEP community are interfaced.
\end{description}

@@

Following the structure of the CMSSW \todo[inline]{Referenz? zumindest github-link?} framework, the \artus framework\footnote{\url{https://github.com/artus-analysis/artus}} provides a front-to-end solution for the analysis of event-based data.
The basic concept of modules processing events within an event loop is complemented by tools for the configuration of every step.
The complete structure is explained in the next section~\ref{section_artus_structure}.

@@

The basic concept of the framework is taken from the code used for the $\PZ+\text{jet}$ analysis documented in reference~\cite{joram_phd}.
It is generalized and extended to satisfy the needs of various analysis groups.
The structure is illustrated in figure~\ref{figure_artus_structure}.

@@

The events are then sequentially processed by processors organized in pipelines.
First, a so-called global pipeline is run.
Then, the execution is split up into multiple so-called local pipelines, that are also processed sequentially.
The composition of a pipeline is illustrated in figure~\ref{figure_artus_pipeline}.

@@

\todo[inline]{I would not put the remaining part of this chapter here.
It should be either part of the requirements or part of the summary what Artus does achieve; but in this technical part in between it feels misplaced
-> Should we put this part into a dedicated section directly after this one?
I think the idea was to explain how certain design decisions, e.g.\ pipelines, achieve the \artus design goals.
Therefore it should not be too far from the technical descriptions.
}
This structure extends the established concepts of existing frameworks like FWLite from CMSSW to the following four main advantages:
\begin{description}
\item[Modular structure:] Analysis steps are implemented in processors, breaking down the complete analysis in well-defined and easily understandable and writable parts.
\item[Re-utilization of code:] Processors can be shared among different analysis groups.
Every user contributes to the testing of existing code and improves the confidence in this code.
\item[Configurability:] Each module uses settings that are read from the the configuration.
Each pipeline is mapped to a separate sub-set of the configuration.
\item[High performance:] The concept of pipelines allows to analyze the data with different configurations at the same time, while the data is only read in only once and common processing steps are shared in the global pipeline.
Therefore, the cost-intensive input operations are reduced to a minimum.
\end{description}

@@

The most important part of an analysis are the processors.
The abstract processors of the \artus core need to be implemented to realize the physics needs of the analysis in the form of code.
New processors have to be registered in the factory.
After the basic structure of an analysis is set up, most changes and additions are only related to the processors.

@@

\section{HarryPlotter -- Python Post-processing Framework}
\label{section_artus_harryplotter}

@@

Table~\ref{table_artus_runtime_comparison} shows a runtime measurement for the baseline part of the $\PH\to\Pgt\Pgt$ in the four main channels ($\Pgm\Pgt$, $\Pe\Pgt$, $\Pe\Pgm$ and $\Pgt\Pgt$).
It illustrates the main advantage of the \artus structure: sub-analyses with different sets of configurations can be performed in the same run avoiding the multiple IO overhead when the configurations would be performed sequentially in different runs.

@@

The \artus analysis framework presented in this paper offers a comprehensive and well-integrated solution for event-based data analyses in the HEP domain.
In conjunction with the plotting toolkit HarryPlotter, all data analysis needs are covered.

>>>>

\begin{abstract}
\artus is an event-based data-processing framework for high energy physics experiments.
It is designed for large-scale data analysis in a collaborative environment.
The architecture design choices take into account typical challenges and are based on experiences with similar applications.
The structure of the framework and its advantages are described.
An example use case and performance measurements are presented.
The framework is well-tested and successfully used by several analysis groups.
\end{abstract}

@@

\section{Typical Workflow of High Energy Physics Analyses}
\label{section_artus_analysis_workflow}
% This section should contain a short description of a HEP/CMS-typical workflow.
% Especially common problems and challenges should be highlighted.
% From these problems, the design requirements for Artus are derived.
The workflow of the experiments collecting data at the Large Hadron Collider (LHC) at CERN is taken as an example to illustrate the organization and needs of a modern High Energy Physics (HEP) analysis.
Still, equal or very similar requirements are also present for HEP experiments not located at the LHC.

Between the data taking at the detector or the generation of simulated events with the subsequent detector simulation and the presentation of the physical results, the following analysis steps have shown to be the best compromise in terms of turnaround time and efficiency as explained in Figure \ref{figure_analysis_steps}.

The reconstruction of the raw data coming from the experiments is performed on large computing centers organized in the WLCG\footnote{\url{http://wlcg-public.web.cern.ch}}, the Worldwide LHC Computing Grid.
The resulting datasets contain event information and high-level physics objects and are of $\mathcal O(10-100)$~TB in size.
\todo[inline]{move file size numbers to Figure \ref{figure_analysis_steps}?}

@@

Source code and configuration files have become the most precise form of description.
Therefore, readability also for people who have not written the code is crucial.
Also, if possible, existing and well-established libraries of the HEP community are interfaced.
\end{description}

@@

Following the structure of the CMSSW~\cite{cmssw} \todo[inline]{Referenz? zumindest github-link?} framework, the \artus framework\footnote{\url{https://github.com/artus-analysis/artus}} provides a front-to-end solution for the analysis of event-based data.
The basic concept of modules processing events within an event loop is complemented by tools for the configuration of every step.
The complete structure is explained in the next section~\ref{section_artus_structure}.

@@

The basic concept of the framework is taken from the code used for the $\PZ+\text{jet}$ analysis documented in reference~\cite{joram_phd}.
It is generalized and extended to satisfy the needs of various analysis groups.
The structure is illustrated in Figure~\ref{figure_artus_structure}.

@@

The events are then sequentially processed by processors organized in pipelines.
First, a so-called global pipeline is run.
Then, the execution is split up into multiple so-called local pipelines, that are also processed sequentially.
The composition of a pipeline is illustrated in Figure~\ref{figure_artus_pipeline}.

@@

\todo[inline]{I would not put the remaining part of this chapter here.
It should be either part of the requirements or part of the summary what Artus does achieve; but in this technical part in between it feels misplaced
-> Should we put this part into a dedicated section directly after this one?
I think the idea was to explain how certain design decisions, e.g.\ pipelines, achieve the \artus design goals.
Therefore it should not be too far from the technical descriptions.
}
This structure extends the established concepts of existing frameworks like FWLite from CMSSW to the following four main advantages:
\begin{description}
\item[Modular structure:] Analysis steps are implemented in processors, breaking down the complete analysis in well-defined and easily understandable and writable parts.
\item[Re-utilization of code:] Processors can be shared among different analysis groups.
Every user contributes to the testing of existing code and improves the confidence in this code.
\item[Configurability:] Each module uses settings that are read from the the configuration.
Each pipeline is mapped to a separate sub-set of the configuration.
\item[High performance:] The concept of pipelines allows to analyze the data with different configurations at the same time, while the data is read in only once and common processing steps are shared in the global pipeline.
Therefore, the cost-intensive input operations are reduced to a minimum.
\end{description}

@@

The most important part of an analysis are the processors.
The abstract processors of the \artus core need to be implemented to realize the physics needs of the analysis in the form of code.
After the basic structure of an analysis is set up, most changes and additions are only related to the processors.

@@

\section{HarryPlotter -- A Python Framework for Post-processing}
\label{section_artus_harryplotter}

@@

Table~\ref{table_artus_runtime_comparison} shows a runtime measurement for the baseline part of the $\PH\to\Pgt\Pgt$ in the four main decay channels ($\Pgm\Pgt$, $\Pe\Pgt$, $\Pe\Pgm$ and $\Pgt\Pgt$).
It illustrates the main advantage of the \artus structure: sub-analyses with different sets of configurations can be performed in the same run avoiding the multiple IO overhead when the configurations would be performed sequentially in different runs.

@@

The \artus analysis framework presented in this paper offers a comprehensive and well-integrated solution for event-based data analyses in the HEP domain.
In conjunction with the post-processing toolkit HarryPlotter, all data analysis needs are covered.
